### 2017-05-09 paper/introduction.tex
case, abstract

The Vienna Development Method (VDM) is a formal method that supports modeling and analysis of software systems at various levels of abstraction. By the VDM specification language (VDM-SL), one can specify software systems with rich types,