### 2017-05-19 paper/obligations.tex

Objective: explain the semantics of Overture generated POs.

Explain what is proof obligation in VDM. There are kinds of POs: domain, subtype, satisfiability, and termination. Any example?

Explain POs generated by Overture tool. Any example?