### 2017-05-19 paper/related.tex

(1) work about discharging POs using theorem provers. (2) other work about validation/verfication of VDM models rather than discharging POs.

For example, \cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} translate VDM models to HOL and then prove the POs within the environment of HOL. However, this approach has several issues: (1) the translation is applied to the whole VDM model so that it may be tedious and error-prone; (2) The proof process is usually complicated and requires an expert; (3) It is not clear where the problem is when the proof fails.

\cite{AL:97:POGV,CFP:14:TVCSAP,CT:15:EOCGTIS}