### 2017-05-26 paper/conclusions.tex
conslusions complete

In this paper, we have proposed an approach to validation of VDM models by discharging proof obligations (POs) with SMT solvers. In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT sover is used for encoding and solving formulas of generated POs. Encode strategies including steps of encoding as well as patterns of encoding are demonstrated and two case studies are reported. From our experiences, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers when the discharging fails can provide useful guidance for correcting inconsistency of VDM models. We also discussed other issues such as applicability of encoding strategies considering the automation of our approach. For the next steps of our approach, we are considering more patterns of encoding and how to organize the patterns of encoding for challenging the automation issue.