### 2017-05-27 paper/conclusions.tex
polished (all)

In this paper, we have proposed an approach to validation of VDM models by discharging proof obligations (POs) with SMT solvers. In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT sover is used for encoding and solving formulas of generated POs. Encoding strategies including steps of encoding as well as encoding patterns are demonstrated, and two case studies are reported. From our experiences, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers when the discharging fails can provide useful guidance for correcting inconsistency of VDM models. We also discussed other issues such as applicability of encoding strategies considering the automation of our approach. As the future work, we are gathering and organizing more encoding patterns for challenging the automation issue.