### 2017-05-27 paper/related.tex

% Development history of VDM POG.
Proof is essential to guarantee the consistency of VDM models~\cite{978-3-540-19813-0}, and the proof obligation generation of VDM is firstly proposed by B. K. Aichernig and P. G. Larsen~\cite{AL:97:POGV}. Later, A. Ribeiro and P. G. Larsen worked on proof obligation generation and discharging related to termination of recursive functions~\cite{Ribeiro2010}.

% Work related to discharging POs using theorem provers. 
There are several researches focusing on translation from VDM to theorem provers and discharging POs of VDM with theorem provers. S. Agerholm~\cite{Agerholm1996} proposed a translation from a subset of VDM-SL to PVS for type checking. S. Maharaj and J. Bicarregui~\cite{632849} used Agerholm's translation to assist the verification of VDM-SL models and their refinements. Later, S. D. Vermolen et al.~\cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} utilized the parsing mechanism of Overture tools and developed automated translation from a subset of VDM-SL to HOL theorem prover. Vermolen's work also provided some tactics to support automated discharging of POs. L. D. Cuto et al.~\cite{CFP:14:TVCSAP} used the proof obligation generation mechanism of Overture tools for translating CML, a modeling language combing VDM-SL and CSP, to Isabella/HOL for automated proof. Recently, work on improvement of translating VDM-SL to Isabella/HOL are proposed~\cite{CT:15:EOCGTIS}.

% Comparison to our approach
The theorem prover based approaches has common issues: (1) the translation from VDM-SL to theorem provers may be errorneous since the translation applies on the whole VDM-SL model; (2) The proof process is usually complicated and requires an expert even with the help of proof tactics; (3) Theorem provers provide little information when they fail to discharge POs. Compare to our approach that discharges POs of VDM-SL with SMT solvers, we adopt the efficiency of SMT solvers and the encoding does not require to be applied on the whole VDM-SL model. Furthermore, the counterexample generation of SMT solvers is helpful for locating the problem. However, encoding in our approach is manual at this point and still need improvements.

% Work related to validation/verfication of VDM models with model checking.
On the other hand, verification and validation of VDM-SL models with model checking techniques are recently proposed. K. Lausdahl~\cite{kenneth:ifm2013} proposed a semantics-preserving translation that constructs an Alloy model from a subset of VDM-SL. This work aims to support the validation of implicitly specified VDM-SL models with Alloy. H-H. Lin et al.~\cite{DBLP:conf/ftscs/LinOKA15} proposed an approach that combines SPIN model checker and the VDMJ interpreter for verifying explicitly specified VDM-SL models. Lausdahl's work is similar to our approach since Alloy uses SAT solvers, but SMT solvers in our approach can solve wider formulas with more efficiency. Both Lausdahl's and Lin's work require types involved in a VDM-SL model being bounded, and there are issues for Alloy and SPIN to represent the rich and complex VDM-SL types. Compare to them, bounds are not necessary for SMT solvers in our approach and our encoding patterns can applied to complex types thanks for build-in theories in SMT sovers.

% For example, \cite{Verm:2007:master,Vermolen:2010:PCV:1774088.1774608} translate VDM models to HOL and then prove the POs within the environment of HOL. However, this approach has several issues: (1) the translation is applied to the whole VDM model so that it may be tedious and error-prone; (2) The proof process is usually complicated and requires an expert; (3) It is not clear where the problem is when the proof fails.

% \cite{CFP:14:TVCSAP,CT:15:EOCGTIS}