### 2017-07-31 memocode_cr/conclusions.tex
MEMOCODE'17 camera ready

In this paper, we have proposed an approach to validation of VDM models by discharging proof obligations (POs) with SMT solvers. In our approach, Overture tool is used as the proof obligation generator, and the Python API of Z3 SMT solver is used for encoding and solving formulas of generated POs. Encoding strategies including steps of encoding as well as encoding patterns are demonstrated, and two case studies are reported. As initial results, it is recognized that discharging POs using SMT solvers is efficient, and the counterexample generation of SMT solvers is useful for correcting inconsistency of VDM models. Currently the encoding is done by hand. We are working on more concrete encoding patterns and automatic encoding tool for more case studies.
% We also discussed other issues such as the applicability of encoding strategies considering the automation of our approach. As the future work, we are gathering and organizing more encoding patterns for challenging the automation issue.