### 2017-07-31 memocode_cr/strategy.tex

% Explain the encode strategy: (1) explain Z3 and python API; (2) strategy on types; (3) take negation or not

% Z3 introduction, API introduction
In this work, we encode and prove VDM POs with the Z3 SMT solver. Z3 is one of the popular SMT solver wildly used in verification of programs and software systems. Z3 has APIs for major programming languages such as C, Java, and Python. The APIs provide a better way of constructing SMT formulas than SMT-LIB~\cite{BarFT-SMTLIB} because one can adopt the flow control characteristics like loop and if-then-else of programming languages to build and solve formulas in a flexible and smart way. In our approach, we chose Z3's Python API (Z3py) as the encoding environment.

% steps of encoding
% 1. check context information: types, functions, and operations involved in the obligation
% 2. encode context information: types and type invariants, functions, and operations (pre- and post-conditions if needed)
% 3. negation and forall quantifier elimination
Currently, the encoding is manually performed based on the obligation to be discharged and below are the steps of encoding and solving an obligation with SMT solvers.

\begin{enumerate}
\item
Determine the context information of the PO.
% : The context information is the definitions specified in the VDM model that are involved in the obligation. The context information may include types, functions, and operations. Note that if a type is involved, its type invariant should be included, too.
\item
Encode context information.
% : this step encodes the context information determined by the first step and can be divided into several minor steps.
\begin{enumerate}
\item
For a VDM type in the context information, both the type itself and its invariant have to be encoded.
\item
If pre/post-conditions of functions/operations are used in a PO in the form of {\tt pre\_<name>} or {\tt post\_<name>}, encoding of pre/post-conditions is needed.
%  are used in the PO. For example, PO7 of the {\tt CMDS} module described in Section~\ref{section:proof-obligations} mentioned both precondition and postcondition, so both conditions shall be encoded.
\item
If a PO only considers a part of the specification of a type, a function, or an operation, it is only needed to encode the expressions that are directly used in the PO.
%  In other words, if a type, a function, or an operation is not directly mentioned, for example, a variable mentioned with {\tt name:type} style, or a function mentioned with {\tt fun\_name(arguments)} style, there is no need to encode the full specification of the type, the function, or the operation.
\end{enumerate}
\item
Obligation negation and quantifier elimination.
\begin{enumerate}
\item
Obligation negation: In model checking with SMT solvers, the assertion to be checked is usually negated, and the result of unsat means the proof of the assertion succeeds. For the result of sat which means the assertion can be violated, SMT solvers return a model as the counterexample that shows an example of how the assertion is violated. In this work, we also negate the PO to be discharged. However, the elimination of universal quantifier should be considered together when applying the negation to the PO.
\item
Elimination of universal quantifiers: POs of VDM are usually with universal quantifiers. Though Z3 accepts formulas with universal quantifiers, there are limitations applying universal quantifier in Z3. Therefore, we eliminate universal quantifiers in the predicate of a PO before encoding. Obligation negation described above also achieves the elimination of universal quantifiers. 
% On the other hand, for POs with existential quantifiers, obligation negation is not needed for not introducing universal quantifiers. 
However, for POs with both universal and existential quantifiers, or multiple universal quantifiers, it should be avoid introducing troublesome universal quantifiers, i.e., universal quantifiers on sequences, sets, and maps, from exists quantifiers.
\end{enumerate}
\item
Encode the predicate of obligation and check for satisfiability.
\end{enumerate}

% More details of encoding types, operators, and expressions
VDM has rich types from natural number and quotes to sequences and maps. There are no direct matching for all VDM types to the types in Z3. For example, to encode a natural number, we need to declare an integer and then add a constraint of greater and equal to zero because there is no natural number type in Z3. The enumeration type in Z3 can be used to encode simple quote types. For complex types like sequences, maps, or records, we may use uninterpreted functions or arrays as the base type and capture the characteristic and operators of the VDM types by adding corresponding constraints. At this point, there is not yet a complete and systematic way of encoding VDM types in Z3. What we can do is to choose the encoding carefully based on the types and expressions involved in a PO. Based on our case studies, there are patterns of encoding that are frequently used for corresponding expression styles in VDM-SL models.